---
title: "Illustrating the effect of censorship on a QUAIDS system"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setting-up the problem

The following is just uploading the necessary libraries and setting the seed for our simulations

```{r}
set.seed(123)
library(numDeriv) # To compute Hessian and Jacobian
library(mvtnorm)  # To simulate a system of equations
library(BB)       # To solver for the parameters of a system of equations
```

The following creates the dataset. Notice that we are simplifying the problem by assuming that the price index is equal to 1. Thus, our system is no longer non-linear, but it still can reproduce elasticities that follow the Cournot and the Engel aggregations.

```{r}
n <- 10000 # Number of observations

# Generate random prices
lnp1 <- rnorm(n, mean = 1, sd = 1)
lnp2 <- rnorm(n, mean = 2, sd = 2)
lnp3 <- rnorm(n, mean = 3, sd = 3)

lnw <- rnorm(n, mean = 1, sd = 2) # Generate random expenditure

# Parameters and parameter restrictions
alpha1 <- .1
alpha2 <- .2
alpha3 <- 1 - alpha1 - alpha2

beta1 <- .1
beta2 <- .2
beta3 <- - beta1 - beta2

gamma11 <- .1
gamma12 <- .12
gamma13 <- - gamma11 - gamma12

gamma21 <- gamma12
gamma22 <- .22
gamma23 <- - gamma21 - gamma22

gamma31 <- gamma13
gamma32 <- gamma23
gamma33 <- - gamma31 - gamma32


```

For this particular example, I will assume that the variance of the shares 1 and 2 are idential at 2, but they are correlated $\rho = .6$.

```{r}
reducedSgima <- matrix(c(2,0.1,0.1,2), nrow = 2, ncol = 2)
epsilon <- rmvnorm(n, rep(0, 2), sigma = reducedSgima) # Assuming normality
eps1 <- epsilon[,1]
eps2 <- epsilon[,2]
eps3 <- -(eps1 + eps2)

lnpindex <- 0 + alpha1*lnp1 + alpha2*lnp2 + alpha3*lnp3
for(i in as.character(1:3)) {
  for(j in as.character(1:3)) {
    
    gammaij <- get(paste0("gamma", i, j))
    lnpi  <- get(paste0("lnp", i))
    lnpj  <- get(paste0("lnp", j))
    lnpindex <- lnpindex + 0.5*gammaij*lnpi*lnpj
    
  }
}

s1 <- alpha1 + gamma11*lnp1 + gamma12*lnp2 + gamma13*lnp3 + beta1*(lnw - lnpindex) + eps1
s2 <- alpha2 + gamma21*lnp1 + gamma22*lnp2 + gamma23*lnp3 + beta2*(lnw - lnpindex) + eps2
s3 <- alpha3 + gamma31*lnp1 + gamma32*lnp2 + gamma33*lnp3 + beta3*(lnw - lnpindex) + eps3
data <- data.frame(s1, s2, s3, lnp1, lnp2, lnp3, lnw)
```

The following solves for the parameters.

```{r}
param0 <- rep(0, 7)
param0 <- c(param0,
            1,
            0,
            1)

source('/Users/noejnava/Library/CloudStorage/OneDrive-UniversityofIllinois-Urbana/Drive/Projects/censorship_demonstration/code/aids.logLike.R')  

optimOut0 <- BBoptim(par = param0,
                     fn = aids.logLike,
                     s1 = s1,
                     s2 = s2,
                     lnp1 = lnp1,
                     lnp2 = lnp2,
                     lnp3 = lnp3,
                     lnw = lnw,
                     method = 1,
                     lower = c(rep(-Inf, 7), 0, -Inf, 0),
                     upper = rep(Inf, 10),
                     control = list(maximize = TRUE,
                                    M = 1,
                                    trace = TRUE,
                                    maxit = 1000))
```
The following code estimates the elasticities with their respective standard errors using the delta method.

```{r}
source('/Users/noejnava/Library/CloudStorage/OneDrive-UniversityofIllinois-Urbana/Drive/Projects/censorship_demonstration/code/etas.estimation.R')
# Function to calculate etas at their means
etas0 <- etas.estimation(params = optimOut0$par[1:7],
                         s1,
                         s2,
                         s3, 
                         lnp1, 
                         lnp2,
                         lnp3)
# Approximation of S.E. via Delta Method
# Derivative of a etas.estimation() function.
# This is R^p -> R^m where p is the number of parameters, and m is etas
der <- jacobian(f = etas.estimation,
                x = optimOut$par[1:7],
                s1 = s1,
                s2 = s2,
                s3 = s3,
                lnp1 = lnp1,
                lnp2 = lnp2,
                lnp3 = lnp3,
                method = "simple")
# Approximate hessian of the aids.logLike() function.
# This is a mxm matrix that includes the sigmas in the Sigma matrix.
hess <- hessian(x = optimOut$par,
                aids.logLike,
                s1 = s1,
                s2 = s2,
                lnp1 = lnp1,
                lnp2 = lnp2,
                lnp3 = lnp3,
                lnw = lnw)
hessReduced <- hess[1:7,1:7]
vcov <- solve(t(hessReduced) %*% hessReduced) # To be used to calculate SE
SE <- sqrt(diag(der %*% vcov %*% t(der)))
SE <- matrix(SE, nrow = length(SE), ncol = 1)
```

A concern with the variables $s_1$, $s_2$, and $s_3$ is that they cannot be observed by the researcher. Instead, the researcher observes $s_m \in [0,1]$. This creates a problem as I now demonstrate.

```{r}
source('/Users/noejnava/Library/CloudStorage/OneDrive-UniversityofIllinois-Urbana/Drive/Projects/censorship_demonstration/code/obs.frm.lat.shr.R')
S <- cbind(s1, s2, s3)
S <- ifelse(S < 0, 0, S)
denominator <- rowSums(S)
S <- S/denominator
s1_obs <- S[,1]
s2_obs <- S[,2]
s3_obs <- S[,3]

optimOut1 <- BBoptim(par = param0,
                     fn = aids.logLike,
                     s1 = s1_obs,
                     s2 = s2_obs,
                     lnp1 = lnp1,
                     lnp2 = lnp2,
                     lnp3 = lnp3,
                     lnw = lnw,
                     method = 1,
                     lower = c(rep(-Inf, 7), 0, -Inf, 0),
                     upper = rep(Inf, 10),
                     control = list(maximize = TRUE,
                                    M = 1,
                                    trace = TRUE,
                                    maxit = 1000))
etas1 <- etas.estimation(params = optimOut1$par[1:7],
                         s1,
                         s2,
                         s3, 
                         lnp1, 
                         lnp2,
                         lnp3)
present.etas(etas0)
present.etas(etas1)
```


